---
title: "cartogram: Create Cartograms with R"
output:
  github_document:
    fig_width: 4
    fig_height: 3.5
---

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/cartogram)](https://cran.r-project.org/package=cartogram)
[![R-CMD-check](https://github.com/sjewo/cartogram/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/sjewo/cartogram/actions/workflows/R-CMD-check.yaml)
[![CRAN Downloads](https://cranlogs.r-pkg.org/badges/cartogram)](https://cran.r-project.org/package=cartogram)
<!-- badges: end -->

```{r, echo=F}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

`cartogram` is an R package that implements methods for generating continuous area cartograms (based on the rubber sheet distortion algorithm by Dougenik et al., 1985), non-contiguous area cartograms (Olson, 1976), and non-overlapping circles cartograms (Dorling et al., 1996).

## Installation

You can install the **cartogram** package from CRAN as follows:

```{r, eval=FALSE}
install.packages("cartogram")
```

To upgrade to the latest development version of `cartogram`, install the package `remotes` and run the following command:

```{r, eval=FALSE}
remotes::install_github("sjewo/cartogram")
```

## Examples

### Continuous Area Cartogram

```{r cont, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)

data("World")

# Keep only the African continent
afr <- World[World$continent == "Africa", ]

# Project the map
afr <- st_transform(afr, 3395)

# Construct continuous area cartogram
afr_cont <- cartogram_cont(afr, "pop_est", itermax = 5)

# Plot the cartogram
tm_shape(afr_cont) +
  tm_polygons("pop_est",
              fill.scale = tm_scale_intervals(style = "jenks")) +
  tm_layout(frame = FALSE,
            legend.position = c("left", "bottom"))
```

### Non-contiguous Area Cartogram

```{r ncont, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)

data("World")

# Keep only the African continent
afr <- World[World$continent == "Africa", ]

# Project the map
afr <- st_transform(afr, 3395)

# Plot the original map boundaries
tm_shape(afr) +
  tm_borders() +
  # Add the the cartogram
  tm_shape(cartogram_ncont(afr, "pop_est")) +
  tm_polygons("pop_est",
              fill.scale = tm_scale_intervals(style = "jenks")) +
  tm_layout(frame = FALSE,
            legend.position = c("left", "bottom"))
```

### Non-Overlapping Circles Cartogram

```{r dorling, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)

data("World")

# Keep only the African continent
afr <- World[World$continent == "Africa", ]

# Project the map
afr <- st_transform(afr, 3395)

# Plot the original map boundaries
tm_shape(afr) +
  tm_borders() +
  # Add the the cartogram
  tm_shape(cartogram_dorling(afr, "pop_est")) +
  tm_polygons("pop_est",
              fill.scale = tm_scale_intervals(style = "jenks")) +
  tm_layout(frame = FALSE,
            legend.position = c("left", "bottom"))
```

## Use multiple CPU cores

```{r parallel, fig.asp = 1.2}
library(cartogram)
library(sf)
library(tmap)
library(future)
library(future.apply)
library(parallelly)
library(progressr)

data("World")

# Keep only the African continent
afr <- World[World$continent == "Africa", ]

# Project the map
afr <- st_transform(afr, 3395)

# Create cartogram using 2 CPU cores on the local machine
# This can speed up computation for larger datasets.
# Set show_progress to TRUE for a progress indicator.
afr_cont <- cartogram_cont(afr, weight = "pop_est",
                            itermax = 5, 
                            n_cpu = 2,
                            show_progress = FALSE)

# Plot the cartogram
tm_shape(afr_cont) +
  tm_polygons("pop_est",
              fill.scale = tm_scale_intervals(style = "jenks")) +
  tm_layout(frame = FALSE,
            legend.position = c("left", "bottom"))
```

## Acknowledgements

The non-contiguous area cartogram and non-overlapping circles cartogram functionalities include major code contributions from @rCarto and @neocarto.

@nowosad contributed to the package by transitioning it to use the `sf` package and by enhancing the documentation, a task further supported by documentation improvements from @oliveroy.

The functionality to utilize multiple CPU cores was contributed by @e-kotov.

## References

This package implements algorithms based on the following seminal works:

* Dorling, D. (1996). Area Cartograms: Their Use and Creation. In Concepts and Techniques in Modern Geography (CATMOG), 59.
* Dougenik, J. A., Chrisman, N. R., & Niemeyer, D. R. (1985). An Algorithm To Construct Continuous Area Cartograms. In The Professional Geographer, 37(1), 75-81.
* Olson, J. M. (1976), Noncontiguous Area Cartograms. The Professional Geographer, 28: 371–380. [doi:10.1111/j.0033-0124.1976.00371.x](https://doi.org/10.1111/j.0033-0124.1976.00371.x)
